---
title: "Lab 5 Homework"
author: "Zabrisky Roland"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Load the superhero data
These are data taken from comic books and assembled by fans. The include a good mix of categorical and continuous data.  Data taken from: https://www.kaggle.com/claudiodavi/superhero-set  

Check out the way I am loading these data. If I know there are NAs, I can take care of them at the beginning. But, we should do this very cautiously. At times it is better to keep the original columns and data intact.  
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
```

## Data tidy
1. Some of the names used in the `superhero_info` data are problematic so you should rename them here.  
```{r}
superhero_info<- rename(superhero_info, name="name", gender="Gender", eye_color= "Eye color", race="Race", hair_colorr="Hair color", height="Height", publisher= "Publisher", skin_color="Skin color", alignment="Alignment", weight="Weight")
```

Yikes! `superhero_powers` has a lot of variables that are poorly named. We need some R superpowers...
```{r}
head(superhero_powers)
```

## `janitor`
The [janitor](https://garthtarr.github.io/meatR/janitor.html) package is your friend. Make sure to install it and then load the library.  
```{r}
library("janitor")
```

The `clean_names` function takes care of everything in one line! Now that's a superpower!
```{r}
superhero_powers <- janitor::clean_names(superhero_powers)
```

## `tabyl`
The `janitor` package has many awesome functions that we will explore. Here is its version of `table` which not only produces counts but also percentages. Very handy! Let's use it to explore the proportion of good guys and bad guys in the `superhero_info` data.  
```{r}
tabyl(superhero_info, alignment)
```

2. Notice that we have some neutral superheros! Who are they?
```{r}
superhero_info %>% 
  filter( alignment== "neutral") %>% 
  select("name")
```

## `superhero_info`
3. Let's say we are only interested in the variables name, alignment, and "race". How would you isolate these variables from `superhero_info`?
```{r}
superhero_info %>% 
  select(alignment, name, race)
```

## Not Human
4. List all of the superheros that are not human.
```{r}
superhero_info %>% 
  select(name, alignment, race) %>% 
  filter(race != "Human")
```

## Good and Evil
5. Let's make two different data frames, one focused on the "good guys" and another focused on the "bad guys".
```{r}
good_guys<- filter(superhero_info, alignment== "good")
bad_guys<- filter(superhero_info, alignment== "bad")
```

6. For the good guys, use the `tabyl` function to summarize their "race".
```{r}
tabyl(good_guys, race)
```

7. Among the good guys, Who are the Asgardians?
```{r}
filter(good_guys, race== "Asgardian")
```

8. Among the bad guys, who are the male humans over 200 inches in height?
```{r}
bad_guys %>% 
  filter(race== "Human", gender== "Male", height > 200)
```

9. OK, so are there more good guys or bad guys that are bald (personal interest)?
_Good news! There are more bald good guys than bald bad guys._
```{r}
jledshair1<- filter(good_guys, hair_colorr== "No Hair")
dim(jledshair1)
```
```{r}
Jledshair2<- filter(bad_guys, hair_colorr== "No Hair")
dim(Jledshair2)
```


10. Let's explore who the really "big" superheros are. In the `superhero_info` data, which have a height over 300 or weight over 450?
```{r}
superhero_info %>% 
  filter(weight> 450| height>300)
```

11. Just to be clear on the `|` operator,  have a look at the superheros over 300 in height...
```{r}
superhero_info %>% 
  filter(height>300)
```

12. ...and the superheros over 450 in weight. Bonus question! Why do we not have 16 rows in question #10?
_The number of rows for #10 is higher than 16 because heroes that appear on both tables do not appear twice on #10._
```{r}
superhero_info %>% 
  filter(weight> 450)
```

## Height to Weight Ratio
13. It's easy to be strong when you are heavy and tall, but who is heavy and short? Which superheros have the highest height to weight ratio?
_These are the superheros with the highest weight to height ratios, with the number one spot going to Giganta._
```{r}
superhero_info %>%
  mutate(height_weight = weight/height) %>% 
  select(name, height, weight, height_weight) %>% 
  arrange(desc(height_weight))
```

## `superhero_powers`
Have a quick look at the `superhero_powers` data frame.  
```{r}
glimpse(superhero_powers)
```

14. How many superheros have a combination of accelerated healing, durability, and super strength?
_There are 97 superheroes with this combination._
```{r}
superhero_powers %>% 
  filter(durability== "TRUE", accelerated_healing== "TRUE", super_strength== "TRUE") %>% 
  select(hero_names)
```



## `kinesis`
15. We are only interested in the superheros that do some kind of "kinesis". How would we isolate them from the `superhero_powers` data?


```{r}
superhero_powers %>% 
  select(hero_names, ends_with("esis")) %>% 
  filter_all(any_vars(.==TRUE))
```

16. Pick your favorite superhero and let's see their powers!
```{r}
superhero_powers %>% 
  filter(hero_names=="Hulk")%>% 
  select_if(all_vars(.=="TRUE"))
```